Component response rate variation underlies the stability of complex systems (revision notes)

### 2020-02-23 notebook/revision_notes.tex
Run the revision notes again

\hypertarget{effects-of-correlations-interactions}{%
\section{Effects of correlations
interactions}\label{effects-of-correlations-interactions}}

@@

The effect of \(\gamma\) does not necessarily result in a circular
eigenvalue spectra of \(\textbf{M}\). In both panels of Figure 1,
\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a,b and 1c,d,
\(d = 0\) and \(d = 1\), respectively, causing the eigenvalue spectra of
\(\textbf{A}\) to differ on the real axis (Figure a,c). Figure panels b
and c show eigenvalue spectra after \(\gamma\) is included such that
\(M = \gamma A\), where elements of \(\gamma\) are drawn from a uniform
distribution \(\gamma \sim \mathcal{U}(0, 2)\). In both cases,
\(\sigma^{2}_{A} =\) 0.05 and \(\sigma^{2}_{\gamma} =\) 0.0031.
Nevertheless, the change in eigenvalue spectra caused by \(\gamma\) can
affect \(\max(\Re(\lambda))\) and therefore system stability.

Stability also decreases with increasing \(\rho\) such that
\(\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right)\).
Hence, when \(\rho\) values are negative, such as when \(\textbf{M}\)
models a predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have
opposing signs, stability is expected to
increase\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. When
elements of \(\gamma\) vary independently, the magnitude of \(\rho\) is
decreased because \(\sigma^{2}_{\gamma}\) increases the variance of
\(M_{ij}\) without affecting the covariance between \(M_{ij}\) and
\(M_{ji}\),

\[\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.\]
Because \(\gamma\) does not affect \(cov(M_{ij}, M_{ji})\), the sign of
\(\rho\) remains unchanged.

\hypertarget{more-down}{%
\section{More down}\label{more-down}}

@@

When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
\(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
(95\% CIs: -75.9050127\(\times 10^{-6}\), 84.5804827\(\times 10^{-6}\)).
Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
clearer picture of the effect of \(\gamma\), it is useful to show the
relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above,
but this time also for how the relationship changes given
\(Var(\gamma)\). That is, given \(\textbf{M} = \textbf{A}\) in addition
to \(\textbf{M} = \gamma\textbf{A}\).

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Tang2014b}{}%
1. Tang, S. \& Allesina, S. Reactivity and stability of large
ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
(2014).

\leavevmode\hypertarget{ref-Allesina2012}{}%
2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
\emph{Nature} \textbf{483,} 205--208 (2012).

\leavevmode\hypertarget{ref-Allesina2015a}{}%
3. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

\leavevmode\hypertarget{ref-Gibbs2017}{}%
4. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
population abundances on the stability of large random ecosystems.
\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
Physics} \textbf{98,} 022410 (2018).

>>>>

\section{Effects of correlations
interactions}\label{effects-of-correlations-interactions}

@@

The effect of \(\gamma\) does not necessarily result in a circular
eigenvalue spectra of \(\textbf{M}\). In both panels of Figure 1,
\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a, b and 1c, d,
\(d = 0\) and \(d = 1\), respectively, causing the eigenvalue spectra of
\(\textbf{A}\) to differ on the real axis (Figure a,c). Figure panels b
and c show eigenvalue spectra after \(\gamma\) is included such that
\(\textbf{M} = \gamma \textbf{A}\), where elements of \(\gamma\) are
drawn from a uniform distribution \(\gamma \sim \mathcal{U}(0, 2)\). In
both cases, \(\sigma^{2}_{A} =\) 0.05 and \(\sigma^{2}_{\gamma} =\)
0.0032. Nevertheless, the change in eigenvalue spectra caused by
\(\gamma\) can affect \(\max(\Re(\lambda))\) and therefore system
stability.

Stability also decreases with increasing \(\rho\) such that
\(\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right)\).
When \(\rho\) values are negative, such as when \(\textbf{M}\) models a
predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have opposing
signs, stability is expected to
increase\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. When
elements of \(\gamma\) vary independently, the magnitude of \(\rho\) is
decreased because \(\sigma^{2}_{\gamma}\) increases the variance of
\(M_{ij}\) without affecting the covariance between \(M_{ij}\) and
\(M_{ji}\),

\[\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.\]

Because \(\gamma\) does not affect \(cov(M_{ij}, M_{ji})\), the sign of
\(\rho\) remains unchanged.

\section{More down}\label{more-down}

@@

When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
\(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
(95\% CIs: -80.284156\(\times 10^{-6}\), 86.1295891\(\times 10^{-6}\)).
Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
clearer picture of the effect of \(\gamma\), it is useful to show the
relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above,
but this time also for how the relationship changes given
\(Var(\gamma)\). That is, given \(\textbf{M} = \textbf{A}\) in addition
to \(\textbf{M} = \gamma\textbf{A}\).

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Tang2014b}{}
1. Tang, S. \& Allesina, S. Reactivity and stability of large
ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
(2014).

\hypertarget{ref-Allesina2012}{}
2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
\emph{Nature} \textbf{483,} 205--208 (2012).

\hypertarget{ref-Allesina2015a}{}
3. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

\hypertarget{ref-Gibbs2017}{}
4. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
population abundances on the stability of large random ecosystems.
\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
Physics} \textbf{98,} 022410 (2018).
